---
title: "Midterm Review Lab"
format: html
---



# Setup

```{r}
#| label: setup
#| include: false

library(rio)
library(tidyverse)
options(scipen = 999)

penguins <- import("data/penguins.csv")
```



# Penguin Questions

## Data

**Description:** Size measurements for adult foraging penguins near Palmer
Station, Antarctica. Includes measurements for penguin species, island in Palmer
Archipelago, size (flipper length, body mass, bill dimensions), and sex.

**Format:** 

- A tibble with 344 rows and 8 variables:
- **species:** a factor denoting penguin species (Adélie, Chinstrap and Gentoo)
- **island:** a factor denoting island in Palmer Archipelago, Antarctica (Biscoe, Dream or Torgersen)
- **bill_length_mm:** a number denoting bill length (millimeters)
- **bill_depth_mm:** a number denoting bill depth (millimeters)
- **flipper_length_mm:** an integer denoting flipper length (millimeters)
- **body_mass_g:** an integer denoting body mass (grams)
- **sex:** a factor denoting penguin sex (female, male)
- **year:** an integer denoting the study year (2007, 2008, or 2009)

## Task 01

- Question: Which species of penguin has the largest average body mass?
- Answer: Gentoo

- Question: What is the average body mass for this species of penguin?
- Answer: 5,076

- Question: What is the unit of measurement for body mass in this data set?
- Answer: Grams as noted in the data documentation above.

```{r}
penguins |>
  group_by(species) |>
  ## If they don't add na.rm = TRUE to their code, they won't get an answer,
  ## because of the missing data.
  summarize(avg_body_mass = mean(body_mass_g, na.rm = TRUE))
```


## Task 02

Draw a density plot of flipper length and a separate density plot of body mass.

- Question: How would you describe the distribution of flipper length compared to body mass? Upload your plots to Canvas to support your answer.
- Answer: Below . . .

Flipper length is clearly bi-modal and is positively skewed. Body mass is possibly tri-modal, although less obviously so and is more positively skewed.

The most important part is for them to talk about bi/tri modality and hopefully mention that at least one is skewed.

```{r}
penguins |> ggplot(aes(x = flipper_length_mm)) + geom_density()
penguins |> ggplot(aes(x = body_mass_g)) + geom_density()
```


## Task 03

- Question: In which year did they study the most penguins?
- Answer: 2009

```{r}
penguins |> count(year)
```


## Task 04

- Question: In which year were the Adelie penguins the heaviest, on average?
- Answer: 2008

```{r}
penguins |>
  filter(species == "Adelie") |>
  group_by(year)  |>
  summarize(avg_body_mass = mean(body_mass_g, na.rm = TRUE))
```


## Task 05

Draw a bar plot showing the number of penguins studied every year by species.

- Question: Using your barplot, tell me which penguin species was most frequently studied in 2007?
- Answer: Adelie

```{r}
# My directions are weird. This is technically what I'm asking for.
penguins |>
  ggplot(aes(x = species)) +
  facet_wrap(~year)

# But I'll accept this.
penguins |>
  filter(year == 2007) |>
  ggplot(aes(x = species)) +
  geom_bar()
```


## Task 06

Draw a plot showing the relationship between body mass (x) and flipper length (y). Extra credit (5) if you can add the regression line to your plot.

- Question:  How would you describe this relationship? Upload your plot to Canvas to support your answer.
- Answer: 

The relationship is positive. Give them some extra credit if they mention Loess Curve or local regression. https://en.wikipedia.org/wiki/Local_regression Heck, give them some extra credit if they upload a loess plot, even if they don't really understand it, just for trying/exploring.

Helpful Hint:

- Use geom_point for the scatter plot.
- Use geom_smooth for the regression line.
  - Be sure to tell R you want a linear regression.
  - `geom_smooth(method = lm)`
  - But try it with/without `method = lm`.
  - What do you think it is doing if you don't specify a linear regression?

```{r}
# Linear Regression
ggplot(penguins, aes(x = body_mass_g, y = flipper_length_mm)) +
  geom_point() +
  geom_smooth(method = lm)

# Loess Regression
ggplot(penguins, aes(x = body_mass_g, y = flipper_length_mm)) +
  geom_point() +
  geom_smooth()
```



# Class Survey

Remember that survey I asked you to complete the first week of class? Well, it's back! If you look in the `data` folder, there are surveys for 2024 and 2025. Just imagine all the weird and spurious correlations we could find in this data!

```{r}
survey <- import("data/class_survey_2025.csv")
```

The grain of this data is student/survey response. Each row is a unique survey response and since none of you could take it more than once, each row is also a unique student.

Let's take a quick look at the data:

```{r}
survey
```

I'm sure you remember the ridiculous questions, but what you don't know is how I mapped those questions to column names. So here's a little cheat-sheet for what column name goes with what question.

```{r}
import("data/questions.csv")
```


## Task 07

- Questions: How many students threatened to drop this class because I asked them if they like math?
- Answer: 11

```{r}
survey |> count(like_math)

## Or I guess this works too, and then you see how many rwos there are.
survey |> filter(like_math == "Yes")
```


## Task 08

- Question: Can you easily use the column `used_data_to_make_decision` to determine how many of your classmates have used data to make a decision? 
- Answer: No

Almost no two answers are the same. You would have to read each answer and put the student's answer into a countable format.

```{r}
survey |>
  count(used_data_to_make_decision)
```


## Task 09

- Question: How is the data in the `used_data_to_make_a_decision` column different than the other columns of data we have used?
- Answer: It is raw data entry and is not from a drop-down or controlled choice. Therefore it is highly complex and not easily counted with out having someone categorize the data for us or use an AI to do it for us.

There is no code for this answer.


## Task 10

- Question: What percentage of students in this class are Public Health majors?
- Answer: 40.0

The goal is to also see if they can figure out which colun to use.

```{r}
survey |>
  count(what_degree) |>
  mutate(p = 100*n/sum(n))
```


## Task 11

- Question: On average, who is taller? Pre-Med or public health students?
- Answer:

- Question: Do you think this observation is likely to be meaningful? Why or why not?
- Answer: No. It is a very small sample size of students from a single school.

Give full credit to any answer similar to the above.

```{r}
survey |>
  group_by(what_degree) |>
  summarize(avg_height = mean(how_tall, na.rm = TRUE))
```


## Task 12

- Question: Of those students who claim to be good at counting, what percentage have previously analyzed data?
- Answer: 71.428%

```{r}
survey |>
  filter(good_at_counting == "Yes") |>
  count(ever_analyzed_data) |>
  mutate(p = 100*n/sum(n))
```
