---
title: "Episode V: The Emperor (Penguin) Strikes Back"
output: 
  html_notebook:
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 3
    number_sections: true
    theme: cerule
---



# Setup

```{r setup, include=FALSE}
options(scipen = 999)
library(modelr)
library(rio)
library(tidyverse)

penguins <- import("data/penguins.csv")
```

## Data

**Description:** Size measurements for adult foraging penguins near Palmer
Station, Antarctica. Includes measurements for penguin species, island in Palmer
Archipelago, size (flipper length, body mass, bill dimensions), and sex.

**Format:** 

- A tibble with 344 rows and 8 variables:
- **species:** a factor denoting penguin species (Adélie, Chinstrap and Gentoo)
- **island:** a factor denoting island in Palmer Archipelago, Antarctica (Biscoe, Dream or Torgersen)
- **bill_length_mm:** a number denoting bill length (millimeters)
- **bill_depth_mm:** a number denoting bill depth (millimeters)
- **flipper_length_mm:** an integer denoting flipper length (millimeters)
- **body_mass_g:** an integer denoting body mass (grams)
- **sex:** a factor denoting penguin sex (female, male)
- **year:** an integer denoting the study year (2007, 2008, or 2009)

# Task 01

Question: Which species of penguin has the largest body mass and what is it?
Answer: 

Helpful Hint: Don't accept NA as an answer. Use `?mean` if you don't remember to handle missing values.

```{r}
## YOUR CODE HERE!
```

# Task 02

Question: Draw a distribution plot of flipper length and body mass. How would you describe the distribution of each? Upload your plots to Canvas to support your answer.
Answer: 

```{r}
## YOUR CODE HERE!
```

# Task 03

Question: In which year did they study the most penguins?
Answer:

Helpful Hint:
- Use dplyr verbs so you get an exact answer.

```{r}
## YOUR CODE HERE!
```

# Task 04

Question: In which year were the Adelie penguins the heaviest?
Answer: 

```{r}
## YOUR CODE HERE!
```

# Task 05

Question: Draw a bar plot of penguins studied every year. Be sure to include species and tell me which was the most frequently studied species in 2007.
Answer:

Helpful Hint:

- You can solve this geom_bar() OR geom_col().
- If your plots look stacked, try `position = "dodge"`. It makes life easier.
- Or you can use facet_wrap(). Your call really.

```{r}
## YOUR CODE HERE!
```


# Task 06

Question: Draw a plot showing the relationship between body mass (x) and flipper length (y). How would you describe this relationship? Upload your plot to Canvas. Extra credit (5) if you can add the regression line to your plot. Upload your plot to Canvas to support your answer.
Answer: 

Helpful Hint:

- Use geom_point here.

```{r}
## YOUR CODE HERE!
```

# Task 07

Question: Build a linear regression where flipper length (y) is explained as a function of body mass (x). What is the value of the intercept? What is the coefficient of body mass?
Answer: 

Helpful Hint: 
- This is the formal regression of the plot from Task 04. So if your numbers are obviously wildly different, you did something wrong!!!!

```{r}
## YOUR CODE HERE!
```


# Task 08

> All models are wrong but some are useful.
> -- Aphorism of George Box

Based on your plot from Task 04 and what you can see in the `summary()` of your 
model, is model useful or does it contain tigers? Is it flawed like the first model from the lecture which explained less than 10 percent of the variance of charges?



# Task 09

You can use the `cor.test()` function to calculate the correlation between two columns of data. Correlation describes the "degree of relationship" between two variables in a single summary statistic.

Correlation values ranges from -1 to +1. Positive values of correlation, such as in our finely feathered example, indicate that as one variable increase the other variable increases as well. Negative values of correlation indicate that as one variable increases the other variable decreases.
For a nice synopsis of correlation, see https://statistics.laerd.com/statistical-guides/pearson-correlation-coefficient-statistical-guide.php

Although I understand the desire to summarize the relationship between two sets of numbers in a single summary statistic, it doesn't work. Yes, it is easy to imagine what a correlation value of 1 or -1 means. But all values in-between are . . . simply there. Take for example, the following code, which will calculate for you the correlation between body_mass_g and flipper_length_mm.
```{r}
cor.test(x = penguins$body_mass_g, y = penguins$flipper_length_mm)
```

Although not 1, it is still quite high. But what does that mean? How could we use it? Although the statistical concept of correlation is extremely important and I think we have proven beyond a feather of a doubt the correlation between penguin body mass and flipper length, I do not believe this summary statistic has much practical value. This is especially true compared to the rich insight created by a scatter plot or the wealth of details provided by the coefficients of a linear model.

In Canvas, tell me the correlation summary statistic calculated by the above function.