---
title: "Midterm Review Lab"
format: html
---



# Setup

```{r}
#| label: setup
#| include: false

library(rio)
library(tidyverse)
options(scipen = 999)

penguins <- import("data/penguins.csv")
```



# Penguin Questions

## Data

**Description:** Size measurements for adult foraging penguins near Palmer
Station, Antarctica. Includes measurements for penguin species, island in Palmer
Archipelago, size (flipper length, body mass, bill dimensions), and sex.

**Format:** 

- A tibble with 344 rows and 8 variables:
- **species:** a factor denoting penguin species (Adélie, Chinstrap and Gentoo)
- **island:** a factor denoting island in Palmer Archipelago, Antarctica (Biscoe, Dream or Torgersen)
- **bill_length_mm:** a number denoting bill length (millimeters)
- **bill_depth_mm:** a number denoting bill depth (millimeters)
- **flipper_length_mm:** an integer denoting flipper length (millimeters)
- **body_mass_g:** an integer denoting body mass (grams)
- **sex:** a factor denoting penguin sex (female, male)
- **year:** an integer denoting the study year (2007, 2008, or 2009)

The grain of this data is one row per bird studied. Note that this study took place over three consecutive years.

## Task 01

- Question: Which species of penguin has the largest average body mass? and what is it?
- Answer: 

- Question: What is the average body mass for this species of penguin?
- Answer: 

- Question: What is the unit of measurement for body mass in this data set?
- Answer: 

Helpful Hint: Don't accept NA as an answer. Use `?mean` if you don't remember to handle missing values.

```{r}
## YOUR CODE HERE!
```

## Task 02

Draw a density plot of flipper length and a separate density plot of body mass.

- Question: How would you describe the distribution of flipper length compared to body mass? Upload your plots to Canvas to support your answer.
- Answer: 

Helpful Hints: 

- Is the data normally distributed, positively skewed, or negatively skewed?
- Is the data bi-modal (two humps in the density plot) or multi-modal (more than 2 humps)?

```{r}
## YOUR CODE HERE!
```

## Task 03

- Question: In which year did they study the most penguins?
- Answer:

Helpful Hint:
- Use dplyr verbs so you get an exact answer.

```{r}
## YOUR CODE HERE!
```

## Task 04

- Question: In which year were the Adelie penguins the heaviest, on average?
- Answer: 

```{r}
## YOUR CODE HERE!
```

## Task 05

Draw a bar plot showing the number of penguins studied every year by species.

- Question: Which penguin species was most frequently studied in 2007?
- Answer:

Helpful Hint:

- You can solve this geom_bar() OR geom_col().
- If your plots look stacked, try `position = "dodge"`. It makes life easier.
- Or you can use facet_wrap(). Your call really.
- You can confirm your results with dplyr verbs.

```{r}
## YOUR CODE HERE!
```


## Task 06

Draw a plot showing the relationship between body mass (x) and flipper length (y). Extra credit (5) if you can add the regression line to your plot.

- Question:  How would you describe this relationship? Upload your plot to Canvas to support your answer.
- Answer: 

Helpful Hint:

- Use geom_point for the scatter plot.
- Use geom_smooth for the regression line.
  - Be sure to tell R you want a linear regression.
  - `geom_smooth(method = "lm")`
  - But try it with/without `method = "lm"`.
  - What do you think it is doing if you don't specify a linear regression?

```{r}
## YOUR CODE HERE!
```



# Class Survey

Remember that survey I asked you to complete the first week of class? Well, it's back!

```{r}
survey <- import("data/class_survey.csv")
```

The grain of this data is student/survey response. Each row is a unique survey response and since none of you could take it more than once, each row is also a unique person.

Let's take a quick look at the data.

```{r}
survey
```

## Task 07

- Questions: How many students threatened to drop this class because I asked them if they like math?
- Answer:

Helpful Hint: Use the column, like_math.

```{r}
## YOUR CODE HERE!
```

## Task 08

- Question: Based on these survey responses, can you (easily) use the skills we have learned in this class to determine how many of your classmates have used data to make a decision?
- Answer: 

Helpful Hints: 
- Use the column, used_data_to_make_decision
- Valid answers are No, and Yes

```{r}
## YOUR CODE HERE!
```

## Task 09

- Question: How is the data in the used_data_to_make_a_decision column different than the other columns of data we have used?
- Answer:

Helpful Hint: You do not need to write any code to answer this question. See Task 08.


## Task 10

- Question: What percentage of students in this class are Public Health majors?
- Answer: 

Helpful Hint: Use the column degree_program

```{r}
## YOUR CODE HERE!
```


## Task 11

- Question: Of those students who claim to be good at counting, what percentage have previously analyzed data?
- Answer:

Helpful Hint: Use the columns good_at_math and ever_analyzed_data to answer this question.

```{r}
## YOUR CODE HERE!
```